---
title: "10_correlations"
author: "Nick Schultz"
date: "3/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(tidyverse)
library(broom)
library(Hmisc)
```
#### Read the data
```{r}
mh <- read_csv("MillerHadenData.csv")
```
#### Look at the data
```{r}
glimpse(mh)
head(mh)
tail(mh)
mh
```
#### Assumptions:  
* Is data interval/ratio/ordinal?  
* Is there a data point for each participant on both variables?  
* Is data normally distributed?  
* Does the realationship appear linear?  
* Does the spread have homoscedasticity?  

Spearman correlation is non-parametric and can run with interval/ordinal/ratio data.

```{r}
is.na(mh)

ggplot(mh, aes(Abil)) +
  geom_histogram(bins = 30)

ggplot(mh, aes(IQ)) +
  geom_histogram()

ggplot(mh, aes(x = IQ, y = Abil)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  theme_minimal()
```

```{r}
results <- cor.test(mh$Abil, mh$IQ, method = "pearson", alternative = "two.sided") %>%
  tidy()

pvalue <- results %>%
  pull(p.value) %>%
  round(3)

df <- results %>%
  pull(parameter)

correlation <- results %>%
  pull(estimate) %>%
  round(2)
```

#### Write up  
* A pearson correlation found reading ablility and intelligence to be positively correlated with a medium to strong relationship, (r(`r df`) = `r correlation`, p = `r pvalue`). As such we can say that our hypothesis is supported and that there appears to be a relationship between reading ability and IQ in that as reading ability increases so does intelligence.

```{r}
pairs(mh)

mh_mx <- as.matrix(mh)
rcorr(mh_mx, type = "pearson")
```

```{r}
vaping <- read_csv("VapingData.csv")

glimpse(vaping)
sum(is.na(vaping$IAT_BLOCK5_Acc))
head(vaping)

```

```{r}
vaping2 <-  vaping %>%
  filter(IAT_BLOCK3_Acc <= 1 & IAT_BLOCK5_Acc <=1) %>%
  mutate(IAT_average_acc = (IAT_BLOCK3_Acc + IAT_BLOCK5_Acc)/2) %>%
  filter(IAT_average_acc >= .8) %>%
  mutate(IAT_RT = IAT_BLOCK5_RT - IAT_BLOCK3_RT)

length(vaping$IAT_BLOCK3_Acc)
length(vaping$IAT_BLOCK5_Acc)

length(vaping2$IAT_BLOCK3_Acc)
length(vaping2$IAT_BLOCK5_Acc)

head(vaping2)
```

```{r}
descriptives <- vaping2 %>%
  summarise(n = n(),
            mean_IAT_ACC = mean(IAT_average_acc),
            mean_IAT_RT = mean(IAT_RT),
            mean_VPQ = mean(VapingQuestionnaireScore, na.rm = TRUE)) 
```

```{r}
ggplot(vaping2, aes(VapingQuestionnaireScore)) +
  geom_histogram() +
  theme_minimal()

ggplot(vaping2, aes(IAT_RT)) +
  geom_histogram() +
  theme_minimal()

ggplot(vaping2, aes(x = IAT_RT, y = VapingQuestionnaireScore)) +
  geom_point() +
  theme_minimal()

vaping3 <- vaping2 %>%
  filter(!is.na(VapingQuestionnaireScore)) %>%
  filter(!is.na(IAT_RT))
```

```{r}
vap_results <- cor.test(vaping3$VapingQuestionnaireScore,
                        vaping3$IAT_RT,
                        method = "pearson") %>%
  tidy()

correlation <- vap_results %>%
  pull(estimate) %>%
  round(3)

df <- vap_results %>%
  pull(parameter) 

pvalue <- vap_results %>%
  pull(p.value) %>%
  round(3)
```

#### Write up  
* Testing the hypothesis that there would be a relationship between implicit and explicit vaping, a pearson correlation found no significant relationship between IAT reaction times (implicit attitude) and answers on a Vaping Questionnaire (explicit attitude), r(`r df`) = `r correlation`, p = `r pvalue`. Overall this suggests that there is no direct relationship between implicit and explicit attitudes when relating to Vaping and as such our hypothesis was not supported; we cannot reject the null hypothesis.

#### Outliers
```{r}
mh_IQ <- mh %>%
  select(IQ)

mh_IQ_z <- mh_IQ %>%
  mutate(z_scores = (IQ - mean(IQ)) / sd(IQ))

mh_IQ_z %>% filter(abs(z_scores) > 2.5) %>% count() %>% pull(n)
```
#### A different approach using a correlation table  
```{r}
library(corrr)
```

```{r}
mh %>%
  select(-Participant) %>%
  correlate(method = "pearson", quiet = TRUE) %>%
  shave() %>%
  fashion(decimals = 3, leading_zeros = FALSE)
```

#### Comparing correlations  
```{r}
library(cocor)
```

```{r}
compar1 <- cocor.indep.groups()
```

