---
title: "Within-Subjects t-tests"
author: "Nick Schultz"
date: "3/8/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tibble)
library(tidyverse)
library(broom)
```
### Within-Subjects t-test

#### Assumptions of between-subjects t-test:  
* independence  
* equal variances  
* dependent variable is continuous (interval/ratio)  
* dependent variable normally distributed  

#### Assumptions of within-subjects t-test:  
* all participants appear in both groups  
* dependent variable is continuous (interval/ratio)  
* dependent variable normally distributed  

#### Welch's t-test vs. Student's t-test
* Welch's does not assume equal variance
```{r}
my_data <- tibble(my_iv = rep(c("A", "B"), each = 20),
                  my_dv = c(rnorm(n = 20, mean = 0, sd = 1),
                                  rnorm(20, 0, 1)))
t.test(my_dv ~ my_iv, data = my_data, var.equal = TRUE)
t.test(my_dv ~ my_iv, data = my_data, var.equal = FALSE)
```

### Task 2
```{r}
ratings <- read_csv("GuiltJudgements.csv")
```
```{r}
lates <-  ratings %>%
  filter(Evidence == "Late") %>%
  select(Participant, Evidence, "12", "13") %>%
  rename(Twelve = "12", Thirteen = "13") %>%
  gather(Twelve, Thirteen, key = "Timepoint", value = "GuiltRating")
```

```{r}
ggplot(lates, aes(GuiltRating)) +
  geom_histogram(binwidth = 1) +
  facet_wrap(~ Timepoint) +
  xlab("GuiltRating") +
  theme_bw()

ggplot(lates, aes(Timepoint, GuiltRating)) +
  geom_boxplot()

ggplot(lates, aes(Timepoint, GuiltRating)) +
  geom_violin(aes(fill = Timepoint), alpha = .2) +
  geom_boxplot(width = .5) +
  geom_jitter(aes(fill = Timepoint), width = .1, alpha = .2)
  scale_y_continuous(breaks = c(1:9)) +
  coord_cartesian(ylim = c(1, 9), expand = TRUE) +
  theme_bw()
  
  lates %>%
    spread(Timepoint, GuiltRating) %>%
    mutate(diff = Thirteen - Twelve) %>%
    ggplot(aes(Evidence, diff)) +
    geom_violin() +
    geom_boxplot(fill = "red",
                 width = .5,
                 alpha = .1,
                 outlier.color = "red") +
    geom_jitter(color = "blue",
                width = .1,
                shape = 4) +
    theme_classic()
```

```{r}
descriptives <- lates %>%
  group_by(Timepoint) %>%
  summarise(n = n(),
            mean = mean(GuiltRating),
            sd = sd(GuiltRating),
            se = sd / sqrt(n),
            LowerCI = mean - 1.96 * se,
            UpperCI = mean + 1.96 * se)

  knitr::kable(descriptives, align = "c", caption = "Descriptives data for current study")

ggplot(descriptives, aes(Timepoint, mean, fill = Timepoint)) +
  geom_col(color = "black") +
  geom_errorbar(aes(ymin = LowerCI, ymax = UpperCI),
                position = "dodge", width = .15)
```

```{r}
t.test(GuiltRating ~ Timepoint, lates, paired = TRUE, alternative  = "two.sided") %>% tidy()
```

